Resilience and Home-Space for WSTS

### 2024-01-11 diapo/presentation.tex
up

@@

\phantom{ {resilience},  {$k$-resilience} and {bounded resilience} are not satisfied }

@@

\phantom{ {resilience} hold, {bounded resilience} does not hold }

  \end{frame}
%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
  \begin{frame}{Resilience(s)}

@@

 {\small

{\sc Resilience problems}

\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, and a set $\Safe \subseteq S$.}

@@

\hspace{-0.5cm}  ({\sc resilience problem}) $\Bad \rightarrow^{*} \Safe$ ?

\hspace{-0.5cm}  ({\sc $k$-resilience problem})		$\Bad \rightarrow^{\leq k} \Safe$ ?

\hspace{-0.5cm}  ({\sc bounded resilience problem})	$\exists k ~ S \rightarrow^{\leq k} \Safe$ ?\newline

@@

@@

{resilience}, 
{$k$-resilience} and {bounded resilience} are not satisfied

@@

 {resilience} hold, {bounded resilience} does not hold

@@

  $\Safe_2 = \{q_1(0)\} $
   {bounded resilience} hold, $6$-resilience hold

  \end{frame}
%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
  \begin{frame}{Resilience(s)}

  \hspace{-0.2cm}
{\sc \textcolor{red}{State} resilience problems}

\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, \textcolor{red}{$s \in S$}, and a set $\Safe \subseteq S$.}

@@

\hspace{-0.5cm}{ ({\sc state-resilience problem}) $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$ ?

\hspace{-0.5cm}  ({\sc $k$-state-resilience problem})		$\cancel{\Bad}~ \textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$ ?

\hspace{-0.5cm} ({\sc bounded-state-resilience problem})	$\exists k ~ \cancel{\Bad}~\textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$?\newline}

@@

  \begin{definition}[\cite{DBLP:journals/lmcs/BlondinFM17}]
A {\em Well-Behaved Transition System} (WBTS) 
is a FAC$^1$-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
%the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
%$s_1, t_1 , s_2 \in S$
%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
%	$t_2 \in S$ with 
%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

\phantom{line}

{\small $^1${FAC = all anti-chains are finite}}

    \end{frame}
%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
  \begin{frame}
% TODO
%  \todo[inline,color=blue!20]{ {\it À l'oral~} 
% Resilience(s) "obviously"  undecidable in general so restriction $\to$ restriction to WSTS
% $\to$ so we need to define WSTS properly }

\begin{definition}[\cite{DBLP:journals/iandc/Finkel90}]
A {\em Well-Structured Transition System} (WSTS) 
is a well-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
%the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
%$s_1, t_1 , s_2 \in S$
%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
%	$t_2 \in S$ with 
%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

\begin{block}{Theorem~4}
WBTS $\mathscr{S}=(S,\rightarrow, \leq)$ 
\begin{itemize}
\item $\post$-ideal-effective =effective + $\mathop{\downarrow} \post(I)$ computable %$\post$-ideal-effective 
\item \textcolor{red}{strong} compatibility
\item $\Safe = \mathop{\uparrow} \Safe$
\end{itemize}
{\sc Resilience}, {\sc Bounded resilience} 
and {\sc $k$-resilience} decidable.
\end{block}

@@

@@

 \begin{definition}[\cite{BFM-ic17}]
The \emph{completion}   of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where 
%$Ideals(S)$ is the set of ideals of $S$ and
 $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\mathop{\downarrow} \post_{\mathscr{S}}(I)$. 
% The completion is always \emph{finitely branching} but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo. 
%It is proved in  \cite{BFM-ic17} that $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS. \\
\end{definition}

@@

\begin{exampleblock}{Relation with paths in the completion}
If $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
\end{exampleblock}

@@

\begin{exampleblock}{Relation with paths in the completion}
If $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$. If $\mathscr{S}$ has strong compatibility then $k’ = k$.
\end{exampleblock}

@@

\begin{theorem}\label{SL VASS}
{\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
\end{theorem}

@@

\begin{theorem}\label{vass down}
{\sc Bounded resilience} and {\sc $k$-resilience } are decidable for VASS when 
$\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}

>>>>

\begin{columns}[T]
 \column{0.6\textwidth}

@@

\vspace{0.1cm}

@@

 \column{0.4\textwidth}

\phantom{ {resilience} }

\vspace{0.1cm}

\phantom{ {resilience} }

  \end{frame}
%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
  \begin{frame}{Resilience problems}

@@

\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, and a set $\Safe \subseteq S$, \\ $\Bad$ complement of $\Safe$.}

@@

\begin{columns}[T]
 \column{0.3\textwidth}

  $\Bad \rightarrow^{*} \Safe$  ?

\vspace{0.15cm}

   $\Bad \rightarrow^{\leq k} \Safe$  ?

\vspace{0.15cm}

  $\exists k ~ \Bad \rightarrow^{\leq k} \Safe$   ?\newline

 \column{0.7\textwidth}

\hspace{0.6cm} {\sc  \phantom{bounded} resilience problem}

\vspace{0.16cm}

\hspace{0.6cm} {\sc \phantom{boundi} $k$-resilience problem}

\vspace{0.175cm}

\hspace{0.6cm} {\sc bounded resilience problem}

@@

\only<1>{
\begin{columns}[T]
 \column{0.6\textwidth}

@@

 \column{0.4\textwidth}

 Resilience(s) not satisfied

\vspace{0.1cm}

\phantom{bla} $q_1(m) ~ \cancel{\to} ~ q_0(n)$

\only<2>{
\begin{columns}[T]
 \column{0.6\textwidth}

@@

 \column{0.4\textwidth}

 Resilience hold,

 \vspace{0.1cm}

 \hspace{-0.6cm} Bounded resilience does not

@@

\begin{columns}[T]
 \column{0.6\textwidth}

 \column{0.4\textwidth}

 Bounded resilience hold,

 \vspace{0.1cm}

 \phantom{bla} $6$-resilience hold

$q_1(9) \to q_0(10) \to q_2(0) \to q_1(0)$

  \end{frame}
%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
  \begin{frame}{{State} resilience problems}

\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, and a set $\Safe \subseteq S$, \\ $\Bad$ complement of $\Safe$, \textcolor{red}{$s \in S$}.}

@@

\begin{columns}[T]
 \column{0.45\textwidth}

  $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$  ?

\vspace{0.15cm}

   $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{\leq k} \Safe$  ?

\vspace{0.15cm}

  $\exists k ~ \cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{\leq k} \Safe$   ?\newline

 \column{0.7\textwidth}

\hspace{0.6cm} {\sc  \phantom{bounded} state-resilience problem}

\vspace{0.19cm}

\hspace{0.6cm} {\sc \phantom{boundi} $k$-state-resilience problem}

\vspace{0.20cm}

\hspace{0.6cm} {\sc bounded-state-resilience problem}

@@

  \begin{block}{Definition~\cite{DBLP:journals/lmcs/BlondinFM17}}
A {\em Well-Behaved Transition System} (WBTS) 
is an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
all antichains of $(S, \leq)$ are finite, and
%the transition relation 
$ \rightarrow$ is (upward) compatible with $\leq$. 
%$s_1, t_1 , s_2 \in S$
%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
%	$t_2 \in S$ with 
%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{block}

\begin{block}{Definition~\cite{DBLP:journals/iandc/Finkel90}}
A {\em Well-Structured Transition System} (WSTS) 
is a well-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ with 
 $ \rightarrow$ is (upward) compatible with $\leq$. 
%the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
%$s_1, t_1 , s_2 \in S$
%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
%	$t_2 \in S$ with 
%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{block}

@@

\begin{block}{Theorem~4}
WBTS $\mathscr{S}=(S,\rightarrow, \leq)$ 
\begin{itemize}
\item effective + $\mathop{\downarrow} \post(I)$ computable$^1$ %$\post$-ideal-effective 
\item \textcolor{red}{strong} compatibility
\item $\Safe = \mathop{\uparrow} \Safe$
\end{itemize}
{\sc Resilience}, {\sc Bounded resilience} 
and {\sc $k$-resilience} decidable.
\end{block}

@@

\phantom{Footnote}

{\small $^1$called $\post$-ideal-effective }

@@

 \begin{definition}[\cite{BFM-ic17}]
The \emph{completion}   of a WBTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where 
%$Ideals(S)$ is the set of ideals of $S$ and
 $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\mathop{\downarrow} \post_{\mathscr{S}}(I)$. 
% The completion is always \emph{finitely branching} but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo. 
%It is proved in  \cite{BFM-ic17} that $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS. \\
\end{definition}

@@

\begin{exampleblock}{From a WBTS to its completion}
If $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
\end{exampleblock}

@@

\begin{exampleblock}{From the completion to the WBTS}
If $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$. If $\mathscr{S}$ has strong compatibility then $k’ = k$.
\end{exampleblock}

@@

\begin{block}{Theorem}
{\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
\end{block}

@@

\begin{block}{Theorem}
{\sc Bounded resilience} and {\sc $k$-resilience } are decidable for VASS when 
$\Safe = \mathop{\downarrow} \Safe$.
\end{block}
